Horizontal and vertical diversity jointly drive total biomass and its temporal stability in fish communities

### 2020-06-24 manuscript/bef_stability/manuscript.Rmd
Version sent to coauthors

---
title: Horizontal and vertical diversity drive total biomass and its temporal stability in fish communities
author: Alain Danet, Maud Mouchet, Elisa Thebault and Colin Fontaine
date: \today 
output:
  bookdown::pdf_document2:
    fig_caption: true 
    keep_tex: true
fontsize: 12pt
header-includes:
   - \usepackage{setspace}
   - \doublespacing
   - \usepackage{lineno}
   - \linenumbers
geometry: margin=2cm
bibliography: references.bib
---

@@

@@

@@

Current biodiversity losses and global changes affect ecosystems in complex and
unpredictable ways [@bellard_impacts_2012]. Since the seminal of paper of
@may_will_1972, the relationships between biodiversity and ecosystem properties
have received a lot of attention [@hector_plant_1999; @loreau_biodiversity_2001;
@hooper_effects_2005; @dunne_network_2006; @duffy_why_2009;
@isbell_forest_quantifying_2018]. Biodiversity largely determines both the
amount and the distribution of matter and energy fluxes within ecosystems, and
has hence a tremendous effect on ecosystem properties [@duffy_why_2009], such as
its total biomass [@hector_plant_1999; @tilman_biodiversity_2014;
@grace_integrative_2016; @pennekamp_biodiversity_2018]], biomass production
[@tilman_productivity_1996; @cardinale_species_2002; @grace_integrative_2016]
and the biomass temporal stability [@thebault_trophic_2005;
@thebault_relationship_2006; @tilman_biodiversity_2006;
@olivier_independent_nodate].

@@

Biodiversity can be quantified in numerous ways: either as the number of species
(i.e.  species richness) and in number of functional groups, which are usually
defined based on the trophic ecology of species, namely the way they acquire
food [@allan_stream_2007]. Those metrics can be organized into (i) "horizontal
diversity", which captures the overall diversity of components (i.e. species)
present in a community, and (ii) "vertical diversity" which captures diversity
of components across functional groups, such as the amount trophic categories,
often termed guilds or levels [@allan_stream_2007; @loreau_population_2010;
@zhao_horizontal_2019; @wang_biodiversity_2017]. Vertical diversity has been
primarily quantified as food chain length [@loreau_population_2010;
@zhao_horizontal_2019; @wang_biodiversity_2017] but it can also be characterized
by the food web structure [@wang_biodiversity_2017] such as connectance and
average trophic level. Horizontal and vertical diversity are, however, not
independent from each other. Connectance is known to covary negatively with
species richness [@winemiller_must_1989; @dunne_network_2006;
@thebault_stability_2010]. Species diversity also result in modification of
biomass distribution across the food web [@thebault_relationship_2006;
@duffy_ecosystem_2005; @duffy_functional_2007; @schneider_animal_2016].
Hence, it is crucial to study if horizontal and vertical diversity have synergic
or antagonistic effects and how their interplay modulates ecosystem properties.

@@

Horizontal diversity has been shown to promote higher biomass production in a
variety of ecosystems. Horizontal diversity, as species richness, has been
reported to increase primary productivity [@cardinale_species_2002,
@tilman_productivity_1996; @grace_integrative_2016]. Two mechanisms have been
proposed to rise this effect [@loreau_partitioning_2001]. The complementarity in
resource use, rising through niche differentiation and interspecific
facilitation, increase the total amount of resource used
[@loreau_biodiversity_1998; @cardinale_species_2002; @loreau_partitioning_2001].
Positive select effect arise when species richness leads to an higher dominance
of the most productive species [@tilman_plant_1997; @loreau_partitioning_2001].
As for total biomass, empirical studies have generally demonstrated positive
relationships with species richness in plant communities [@hector_plant_1999;
@tilman_biodiversity_2014; but see @grace_integrative_2016] and mesocosms
[@pennekamp_biodiversity_2018]. When adding the vertical diversity, e.g.
consumers species, @gamfelt_species_2005 and @thebault_relationship_2006 have found that total
biomass increased with increasing species diversity at two trophic levels.
But this positive relation seems to strongly depend on the niche differentiation of
the consumer species, i.e. their resource complementarity,
@thebault_relationship_2006. Hence, the total biomass-species richness
relationship seems to be supported by the same mechanisms within consumers and
within plants. @schneider_animal_2016 also showed in a allometric food web model
that species richness should increase total animal biomass.
Vertical diversity also affect total biomass. Theoretical studies showed that food
web properties, such as interaction distribution and strength, affect the
vertical distribution of biomass [@thebault_relationship_2006;
@schneider_animal_2016; @barbier_pyramids_2019], i.e. average trophic level.
More precisely, an increase of animal diversity results in
higher animal biomass as more biomass is transfered from plant to animals
[@schneider_animal_2016; @thebault_food-web_2003]. However, the role of vertical
diversity on the total biomass remains unclear in
natural communities.

@@

Horizontal and vertical diversity also affects the stability of ecosystem
properties. Experimental studies showed consistently that species richness
increases the temporal stability of biomass in grasslands
[@tilman_biodiversity_2006], and mesocosms [@pennekamp_biodiversity_2018]. The
same relation was found in natural settings for temporal biomass stability in
stream fish communities [@franssen_annual_2011] and temporal abundance stability
of birds and butterflies [@olivier_independent_nodate, but not bats].  Two
mechanisms may govern the temporal stability of communities, (1) species
synchrony and (2) variability [@loreau_species_2008;
@thibaut_understanding_2013]. Species synchrony measures the temporal
covariation of species where low synchrony indicates that species temporal
variation compensate each other at the community level, thereby increasing
temporal stability. Species richness was found to decrease
synchrony between species [@loreau_species_2008; @thibaut_understanding_2013;
@bluthgen_land_2016; @olivier_independent_nodate, but @declerck_species_2006].
The species variability is measured as a average coefficient
of variation of abundance/biomass, weighted by species abundance/biomass. Higher species
temporal variability should thus result in lower temporal stability. Positive
[@tilman_biodiversity_2006], neutral [@jiang_different_2009;
@olivier_independent_nodate] and negative [@franssen_annual_2011;
@olivier_independent_nodate] relationships between species variability and
species richness have been found. This conflicting patterns are dependent on
various mechanisms [see @thibaut_understanding_2013]. Species synchrony has been found to be
the main driver of temporal stability [@tilman_biodiversity_2006;
@olivier_independent_nodate] and responsible for the positive relationship
between horizontal diversity and temporal stability of biomass and abundance.

Vertical diversity, captured by network structure, has been at the heart of the
diversity-stability debate and largely explored by theoretical ecology
[@may_will_1972; @angelis_stability_1975; @dunne_network_2006;
@stouffer_compartmentalization_2011; @thebault_stability_2010;
@barbier_pyramids_2019; @shanafelt_david_w._stability_nodate]. As species
interactions introduce inter-dependencies in species dynamics, they have a
substantial impact on stability. In particular, the strength of predation and
degree of omnivory in food webs have been shown to favor stability
[@neutel__2007]. Furthermore, competition strength and niche breath of
consumers, can lead to negative or positive temporal stability-species richness
relationships [@thebault_trophic_2005] depending on the food web configuration
[@neutel__2007].  However, it is important to note that theoretical ecology
generally defines stability based on community equilibrium and its stability to
small independent perturbations [@may_will_1972; @angelis_stability_1975;
@barbier_pyramids_2019; @shanafelt_david_w._stability_nodate] or to the amount
of secondary extinctions following species removal [@duffy_why_2009]. In
practice, these stability metrics hardly relate to empirical temporal stability
[@arnoldi_resilience_2016].

@@

The environment affects organism performance directly, and thereby, species
abundances and total biomass [Add general ref; @grace_integrative_2016] and
temporal stability [@de_boeck_patterns_2017; @hansen_climate_2013;
@bluthgen_land_2016; @oliver_heterogeneous_2010; @friedpetersen_drivers_2020].
For instance, the hydrology of streams determines the composition of fish
communities, and is also related to fish population temporal stability
[@franssen_annual_2011].  Temperature has potentially conflicting effects on
horizontal and vertical diversity, as higher temperature favour primary
production and fluxes of biomass [@dossena__2012; @ogorman__2012], thus
potentially supporting more species and trophic levels [@arim__2007;
@stegen__2012], but selects against larger body sized individuals
[@dossena__2012; @edeline__2013; @hattab__2016], and thereby should reduce
trophic diversity [@hattab__2016]. It is thus unclear whether these temperature
effects translate indirectly into changes in total standing biomass and
stability in natural food webs. The relative strength of direct and indirect
effects of environment on total biomass and its temporal stability remains an
open question.

- TODO Add on ecosystem size (cf elisa)

@@

We used the ONEMA dataset consisting in standardized sampling of fish
communities in french streams [@edeline__2013]. 99 sites were sampled for more
than 10 years spanning on the 1995-2018 period. We inferred
(**XXX**) potential interaction networks, based on body size and food diet.
Horizontal diversity was quantified as species richness while vertical diversity
was as connectance and average trophic level. Environment was characterized by
the stream size, altitude, slope, and the temperature, flow and enrichment. Finally total
biomass was measured as median total biomass over the sampling period while its
temporal stability was quantified as the inverse of the CV.

@@

Stream fish community constitutes a good model to study the joint effects of
horizontal & vertical diversity on total biomass and its temporal stability.
Despite that fish trophic interactions are difficult to assess in nature, a consequent
body of literature supports that they can be inferred through body size and diet
informations [@gravel_inferring_2013; @jennings_weak_2001;
@brose_predator_2019]. Most of the biomass-diversity and stability-diversity
relationships have been assessed in terrestrial ecosystems while there is a
evident need to test such relationships in a variety of ecosystem types. Stream
ecosystems are characterized by the flowing water and depends for a large part
from the input of terrestrial ecosystems. Furthermore, the biotic activity in
stream ecosystems determines for a large part the water quality
[@cummins_structure_1974] and thus are crucial to maintain **human selfish
goals (can do better than that)**.

@@

Fish communities were monitored across stream sections in France on the period 1995-2018
by the French Office French Office of Water and Aquatic Ecosystems (ONEMA) using
electric fishing, resulting in over XX individual fish sampled [see
@edeline__2013]. Two different standardized protocols were used for little and
big streams (detailed in Supplementary methods). For streams shallower than (XX
m), the sampling takes place over the whole width of the streams and is done by
foot whereas in streams deeper than (XX m), the sampling was done over the bank
only and by foot or by boat. The sampling took place either in late spring and/or Autumn
and temporal series were checked to reduce sampling heterogeneity within
stations (detailed in Supplementary methods). Each station have been followed
ten years or more. The temporal stability of biomass, computed as inverse of the
CV, can cover very different dynamics such as decreasing, stable or increasing.
However, the concept of community variability referred to variability around an
equilibrium, which is rarely observed or difficult to characterize in natural
communities. We finally kept only the 99 stations (Figure \@ref(fig:matmet) B)
which had no trends in biomass (method detailed in Supplementary methods).
However, the results were robust to the relaxation of this hypothesis
(Supplementary results).

@@

To characterize communities variability, abundance or biomass are generally
used. When the body size and therefore body mass vary by several orders of
magnitude, one downside of abundance is that we mix individual that have
potentially really different meaning in terms of ecosystem properties such as
biomass. We therefore estimated the body mass of fish individuals according to
their body size. The body mass of individual fishes were then estimated with an
allometric law ($B_i = 0.01 \times l_i^{3.03}$, $B$ being the body mass in grams,
$l$ the body size in millimeters and $i$ the individual)
[@lleonart_removing_2000; @froese_cube_2006].

The species and community biomass were then computed for each sampling event.
As sampling effort, reported as the surface sampled during the sampling event,
varied across protocol type and sampling event, biomass was reported to the
surface sampled (in $m^{2}$). Community and species biomass were then expressed
in $gm^{-2}$.

Total biomass was characterized as the median total biomass over time. Temporal
biomass stability was defined as the inverse of biomass variability. Variability
was computed as Coefficient of Variation (CV) of the biomass ($CV_i = \sigma_i
\mu_i^{-1}$, with $\sigma_i$ and $\mu_i$ being respectively the standard
deviation and the temporal average of the biomass of the station $i$). Then, temporal
stability was equal to $CV^{-1} = \mu \sigma^{-1}$.

To characterize the mecanisms driving temporal stability, we decomposed
the community CV in synchrony and weighted average species CV
components according to @thibaut_understanding_2013:

$$
CV_{com} = \overline{CV_{sp}} \times \sqrt{\phi}
$$

with $CV_{com}$ being the Coefficient of Variation of the community,
$\overline{CV_{sp}}$ the average coefficient of variation of population
pondered by biomass. $\phi$ is the synchrony, which is the ratio the total
community variance ($\sigma^2_{x_T}$) and the sum of population variance
($\sigma_{x_i}$) [@loreau_species_2008].

$$
\overline{CV_{sp}} = \sum_i{\frac{B_i {CV}_i}{\sum_i{B_i}}}
$$

$$
\phi = \frac{\sigma^2_{x_T}}{\sum_i{\sigma_{x_i}}}
$$

## Hozizontal & vertical diversity

@@

Species richness was computed as the total species richness over time in a given
station. To normalize by sampling effort, we reported species richness to the
sum of surface sampled over the time period, it was expressed in number of
species by square meter.

@@

```{r}
myload(metaweb_analysis, dir= mypath("data"))
meta <- metaweb_analysis
#metaweb_analysis$resource
#meta$metaweb[meta$resource, meta$resource]
```

@@

In each network, we computed two network metrics that have been commonly used to
characterize network structure. Connectance which describes the number of
interactions compared to the number of possible links, computed as ($C = L /
N^2$, L and N being
resp. the number of links and of trophic species plus resource nodes). We then
computed the average trophic level, the average trophic level was weighted by
the biomass of each trophic species (excluding resource nodes for which we had
no biomass) ($\overline{T} = \sum_i^N{B_iT_i \times \sum_i^N{B_i}^{-1}}$, $T_i$
and $B_i$ being resp. the trophic level and the biomass of the node $i$).

In resume, we infered binary trophic networks at the scale of trophic species
(i.e. body size classes by species) to take in account ontogenic diet shifts and
that fish-fish trophic interaction depends on body size ratio. On the other
side, the community metrics were defined at the species level, (i.e. species
richness, average species CV and synchrony).

@@

For each fish station, environmental variables were matched with the year of
fish sampling (except for water temperature for which the record began in 2006).
Then, the average temperature, flow, BOD, river width and depth were computed as
well as coefficient of variation ($CV = \sigma \mu^{-1}$). As habitat and
environmental variables are often collinear, a principal component analysis (PCA)
was performed on the scaled variables. We kept in the analysis the two first
principal components based on eigen values distribution (Fig. SXX).
The PCA explained `r tot_var_expl_pca`%
of the variance. An varimax axis rotation was performed to maximize the representation
of the variables by the principal components (Fig. \@ref(fig:pca-rotated)).
Five principal components were used in the analysis, representing respectively
the (1) river size (river width & depth), (2) altitude, slope and decreasing
average temperature, (3) average enrichment (BOD) and CV of flow, (4) river size
CV (CV of river width and depth) and (5) CV of enrichment (CV of BOD). Further
informations about data transformation and filtering are detailed in
supplementary information.

@@

#### SEM is a convenient approach to investigate direct and indirect path

To reveal the direct and indirect effects of environment, horizontal & vertical
diversity on both total biomass and its temporal stability, we built and
analyzed Structural Equation Models (SEM). It is one promising way to go beyond
bivariate relationships and to investigate complex relationships, direct and
indirect, between variables [@grace_structural_2008; @grace_causal_2014;
@lefcheck_piecewisesem_2016; @grace_integrative_2016]. SEM makes possible to set
causal relationships between variables. This step can be critical, as the
estimated parameters can be determined by both the choice of the variables and
the links between them. Therefore, the set of variables and links should be
supported by theory and literature. We built a metamodel which present the set
of variables and the links between them (Figure \@ref(fig:matmet) A,
\@ref(fig:sem-plot) A & B). Basically, we tested the links between environmental
conditions, horizontal & vertical diversity and both total biomass and its
temporal stability (Figure \@ref(fig:matmet) A). We considered that environment
can directly affect horizontal and vertical diversity, as well as total biomass
and its temporal stability. Horizontal diversity, measured as species richness
directly affects vertical diversity, total biomass and its temporal stability.
Food web structure was shown to be constrained by species richness. In turn,
vertical diversity affects total biomass and its temporal stability (Figure
\@ref(fig:matmet) A).

@@

Structural Equation Modeling are able to assess multivariate relationships,
directs and indirects between variables.  The metamodel describes how variables
are interrelated (Figure \@ref(fig:matmet) A). We built separated structural
equation model for total biomass and its temporal variability (Figure
\@ref(fig:sem-plot) A & B). Structural equation modeling was set up according to the
hypothesis described in introduction and detailed in supplementary information.
Structural equation models were based on Gaussian linear models, some variables
were logged to fulfill model assumptions (see residual plots and data
transformation in supplementary
informations). To account for heterogeneity of hydrographic basin, we set it as random effect on the intercept. Moreover, statistical model were checked for the presence of
multicollinearity (Supplementary information). The slopes reported in the main
text are standardized in order to compare their magnitude ($r_\delta = \beta
\frac{\sigma_x}{\sigma_y}$, $\beta$ being the unstandardized slope, $\sigma_x$
and $\sigma_y$ being respectively the standard deviation of the predictive
variable $x$ and of the response variable $y$). In the main text, only the
slopes which had $P$ value inferior to $0.05$ were reported (but see Table SXX).

The structural equation models used for total biomass and its temporal stability
contains respectively 4 and 6 response variables, 35 and 26 direct paths.

#### Computation of indirect effects

>>>>

---
title: Horizontal and vertical diversity jointly drive total biomass and its temporal stability in fish communities
author: Alain Danet, Maud Mouchet, (Willem Bonnaffe ici?), Elisa Thebault and Colin Fontaine
date: \today 
output:
  bookdown::pdf_document2:
    toc: false 
    fig_caption: true 
    keep_tex: true
fontsize: 12pt
header-includes:
   - \usepackage{setspace}
   - \doublespacing
   - \usepackage{lineno}
   - \linenumbers
geometry: margin=2cm
bibliography: references.bib
---

@@

- Author affiliation:
  - Alain Danet:
  - Maud Mouchet:
  - Willem Bonnaffe:
  - Elisa Thebault: 
  - Colin Fontaine:

- Author contributions: AD, CF, MM and ET designed the study. WB designed the
  network inference method and collected the food diet data. AD performed the
  data analysis and wrote the first draft of the manuscript. All authors
  contributed substantially to revisions.

- Data accessibility statement: the ONEMA database is available upon request to 
  eddy.cosson@afbiodiversite.fr and thierry.point@afbiodiversite.fr. The code
  used for the analysis and to generate the manuscript is available on github:
  (alaindanet/fishcom.git)

- Article type: Letters

- Number of words:
  - Abstract: 194
  - Main text: XXXX

- Number of references: XXX
- Number of figures: 2
- Number of tables: 2
- Number of text boxes: 0

- Coresponding author:
  - Alain Danet
  - UMR 7204 CESCO, BP 61; 45, rue Buffon, 75005 Paris, France
  - alain.danet@mnhn.fr 
  - +331XXXXXXXX

@@

The effect of horizontal  and vertical diversity on ecosystem properties has
been central to ecology. It has been at the basis of the predictions of the
effect species richness and  network structure on total biomass, productivity
and stability. However, they have mainly been studied separately.  We used a
dataset of standardized fish sampling in French stream running from 1995 to
2018. We inferred trophic interactions thanks to recorded body size and diet
literature. We then investigate the direct and indirect effects of stream
size and temperature, horizontal and vertical diversity on total biomass and
its temporal stability. We found that horizontal and vertical diversity had
a positive relationship with total biomass while they had respectively a
negative and positive effect on temporal stability. We also found that
horizontal and vertical diversity are not independent and that environment
had an tremendous effect on total biomass and its temporal stability notably
through its effect on horizontal and vertical diversity. We conclude that our
study highlights that the effects horizontal and vertical diversity should not
treated separately and that it opens exciting perspectives on how environment,
horizontal and vertical can jointly determine total biomass in natural
communities.

(194/150 words)

@@

Current biodiversity losses and global changes affect ecosystems in complex and
unpredictable ways [@bellard_impacts_2012]. Since the seminal of paper of
@may_will_1972, the relationships between biodiversity and ecosystem properties
have received a lot of attention [@hector_plant_1999; @loreau_biodiversity_2001;
@hooper_effects_2005; @dunne_network_2006; @duffy_why_2009;
@isbell_forest_quantifying_2018]. Biodiversity largely determines both the
amount and the distribution of matter and energy fluxes within ecosystems, and
has hence a tremendous effect on ecosystem properties [@duffy_why_2009], such as
its total biomass [@hector_plant_1999; @tilman_biodiversity_2014;
@grace_integrative_2016; @pennekamp_biodiversity_2018], biomass production
[@tilman_productivity_1996; @cardinale_species_2002; @grace_integrative_2016]
and the biomass temporal stability [@thebault_trophic_2005;
@thebault_relationship_2006; @tilman_biodiversity_2006;
@olivier_independent_nodate].

@@

Biodiversity can be quantified in numerous ways: either as the number of species
(i.e.  species richness) and in number of functional groups, which are usually
defined based on the trophic ecology of species, namely the way they acquire
food [@allan_stream_2007]. Those metrics can be organized into (i) "horizontal
diversity", which captures the overall diversity of components (i.e. species)
present in a community, and (ii) "vertical diversity" which captures diversity
of components across functional groups, such as the amount trophic categories,
often termed guilds or levels [@allan_stream_2007; @loreau_populations_2010;
@zhao_horizontal_2019; @wang_biodiversity_2016]. Vertical diversity has been
primarily quantified as food chain length [@loreau_populations_2010;
@zhao_horizontal_2019; @wang_biodiversity_2016] but it can also be characterized
by the food web structure [@wang_biodiversity_2016] such as connectance and
average trophic level. Horizontal and vertical diversity are, however, not
independent from each other. Connectance is known to covary negatively with
species richness [@winemiller_must_1989; @dunne_network_2006;
@thebault_stability_2010]. Species diversity also result in modification of
biomass distribution across the food web [@thebault_relationship_2006;
@duffy_ecosystem_2005; @duffy_functional_2007; @schneider_animal_2016].
Hence, it is crucial to study if horizontal and vertical diversity have synergic
or antagonistic effects and how their interplay modulates ecosystem properties.

@@

Horizontal diversity has been shown to promote higher biomass production in a
variety of ecosystems. Horizontal diversity, as species richness, has been
reported to increase primary productivity [@cardinale_species_2002,
@tilman_productivity_1996; @grace_integrative_2016]. Two mechanisms have been
proposed to rise this effect [@loreau_partitioning_2001]. The complementarity in
resource use, rising through niche differentiation and interspecific
facilitation, increase the total amount of resource used
[@loreau_biodiversity_1998; @cardinale_species_2002; @loreau_partitioning_2001].
Positive select effect arise when species richness leads to an higher dominance
of the most productive species [@tilman_plant_1997; @loreau_partitioning_2001].
As for total biomass, empirical studies have generally demonstrated positive
relationships with species richness in plant communities [@hector_plant_1999;
@tilman_biodiversity_2014; but see @grace_integrative_2016] and mesocosms
[@pennekamp_biodiversity_2018]. When adding the vertical diversity, e.g.
consumers species, @gamfeldt_species_2005 and @thebault_relationship_2006 have found that total
biomass increased with increasing species diversity at two trophic levels.
But this positive relation seems to strongly depend on the niche differentiation of
the consumer species, i.e. their resource complementarity,
@thebault_relationship_2006. Hence, the total biomass-species richness
relationship seems to be supported by the same mechanisms within consumers and
within plants. @schneider_animal_2016 also showed in a allometric food web model
that species richness should increase total animal biomass.
Vertical diversity also affect total biomass. Theoretical studies showed that food
web properties, such as interaction distribution and strength, affect the
vertical distribution of biomass [@thebault_relationship_2006;
@schneider_animal_2016; @barbier_pyramids_2019], i.e. average trophic level.
More precisely, an increase of animal diversity results in
higher animal biomass as more biomass is transfered from plant to animals
[@schneider_animal_2016; @thebault_food-web_2003]. However, the role of vertical
diversity on the total biomass remains unclear in
natural communities.

@@

Horizontal and vertical diversity also affects the stability of ecosystem
properties. Experimental studies showed consistently that species richness
increases the temporal stability of biomass in grasslands
[@tilman_biodiversity_2006], and mesocosms [@pennekamp_biodiversity_2018]. The
same relation was found in natural settings for temporal biomass stability in
stream fish communities [@franssen_annual_2011] and temporal abundance stability
of birds and butterflies [@olivier_independent_nodate, but not bats].  Two
mechanisms may govern the temporal stability of communities, (1) species
synchrony and (2) variability [@loreau_species_2008;
@thibaut_understanding_2013]. Species synchrony measures the temporal
covariation of species where low synchrony indicates that species temporal
variation compensate each other at the community level, thereby increasing
temporal stability. Species richness was found to decrease
synchrony between species [@loreau_species_2008; @thibaut_understanding_2013;
@bluthgen_land_2016; @olivier_independent_nodate; but @declerck_species_2006].
The species variability is measured as a average coefficient
of variation of abundance/biomass, weighted by species abundance/biomass. Higher species
temporal variability should thus result in lower temporal stability. Positive
[@tilman_biodiversity_2006], neutral [@jiang_different_2009;
@olivier_independent_nodate] and negative [@franssen_annual_2011;
@olivier_independent_nodate] relationships between species variability and
species richness have been found. This conflicting patterns are dependent on
various mechanisms [see @thibaut_understanding_2013]. Species synchrony has been found to be
the main driver of temporal stability [@tilman_biodiversity_2006;
@olivier_independent_nodate] and responsible for the positive relationship
between horizontal diversity and temporal stability of biomass and abundance.

Vertical diversity, captured by network structure, has been at the heart of the
diversity-stability debate and largely explored by theoretical ecology
[@may_will_1972; @angelis_stability_1975; @dunne_network_2006;
@stouffer_compartmentalization_2011; @thebault_stability_2010;
@barbier_pyramids_2019; @shanafelt_david_w._stability_nodate]. As species
interactions introduce inter-dependencies in species dynamics, they have a
substantial impact on stability. In particular, the strength of predation and
degree of omnivory in food webs have been shown to favor stability
[@neutel_reconciling_2007]. Furthermore, competition strength and niche breath of
consumers, can lead to negative or positive temporal stability-species richness
relationships [@thebault_trophic_2005] depending on the food web configuration
[@neutel_reconciling_2007].  However, it is important to note that theoretical ecology
generally defines stability based on community equilibrium and its stability to
small independent perturbations [@may_will_1972; @angelis_stability_1975;
@barbier_pyramids_2019; @shanafelt_david_w._stability_nodate] or to the amount
of secondary extinctions following species removal [@duffy_why_2009]. In
practice, these stability metrics hardly relate to empirical temporal stability
[@arnoldi_resilience_2016].

@@

The environment affects organism performance directly, and thereby, species
abundances and total biomass [@grace_integrative_2016] and
temporal stability [@de_boeck_patterns_2017; @hansen_climate_2013;
@bluthgen_land_2016; @oliver_heterogeneous_2010; @friedpetersen_drivers_2020].
For instance, the hydrology of streams determines the composition of fish
communities, and is also related to fish population temporal stability
[@franssen_annual_2011]. Ecosystem size can also increase habitat heterogeneity
and so the number of different resource available, which in turn might increase
resource complementary between species and increase total biomass
[@post_ecosystem_2000; @doi_resource_2009]. Temperature has potentially
conflicting effects on horizontal and vertical diversity, as higher temperature
favour primary production and fluxes of biomass [@dossena_warming_2012;
@ogorman_chapter_2012], thus potentially supporting more species and trophic
levels [@arim_relationship_2007; @stegen_evolving_2012], but selects against
larger body sized individuals [@dossena_warming_2012; @edeline_ecological_2013;
@hattab_forecasting_2016], and thereby should reduce trophic diversity
[@hattab_forecasting_2016]. It is thus unclear whether these temperature effects
translate indirectly into changes in total standing biomass and stability in
natural food webs. The relative strength of direct and indirect effects of
environment on total biomass and its temporal stability remains an open
question.

@@

We used the ONEMA dataset consisting in standardized sampling of fish
communities in french streams [@edeline_ecological_2013]. 99 sites were sampled for more
than 10 years spanning on the 1995-2018 period. We inferred
1496 interaction networks, based on body size and food diet.
Horizontal diversity was quantified as species richness while vertical diversity
was as connectance and average trophic level. Environment was characterized by
the stream size, altitude, slope, and the temperature, flow and enrichment. Finally total
biomass was measured as median total biomass over the sampling period while its
temporal stability was quantified as the inverse of the CV.

Stream fish community constitutes a good model to study the joint effects of
horizontal & vertical diversity on total biomass and its temporal stability.
Despite that fish trophic interactions are difficult to assess in nature, a consequent
body of literature supports that they can be inferred through body size and diet
informations [@gravel_inferring_2013; @jennings_weak_2001;
@brose_predator_2019]. Most of the biomass-diversity and stability-diversity
relationships have been assessed in terrestrial ecosystems while there is a
evident need to test such relationships in a variety of ecosystem types. Stream
ecosystems are characterized by the flowing water and depends for a large part
from the input of terrestrial ecosystems. Furthermore, the biotic activity in
stream ecosystems determines for a large part the water quality
[@cummins_structure_1974] and thus are crucial to maintain stream ecosystem
properties.

@@

@@

Fish communities were monitored across stream sections in France on the period 1995-2018
by the French Office French Office of Water and Aquatic Ecosystems (ONEMA) using
electric fishing, resulting in over XX individual fish sampled [see
@edeline_ecological_2013]. Two different standardized protocols were used for little and
big streams (detailed in Supplementary methods). For streams shallower than (XX
m), the sampling takes place over the whole width of the streams and is done by
foot whereas in streams deeper than (XX m), the sampling was done over the bank
only and by foot or by boat. The sampling took place either in late spring and/or Autumn
and temporal series were checked to reduce sampling heterogeneity within
stations (detailed in Supplementary methods). Each station have been followed
ten years or more. The temporal stability of biomass, computed as inverse of the
CV, can cover very different dynamics such as decreasing, stable or increasing.
However, the concept of community variability referred to variability around an
equilibrium, which is rarely observed or difficult to characterize in natural
communities. We finally kept only the 99 stations (Figure \@ref(fig:matmet) B)
which had no trends in biomass (method detailed in Supplementary methods).
However, the results were robust to the relaxation of this hypothesis
(Supplementary results).

@@

To characterize communities variability, abundance or biomass are generally
used. When the body size and therefore body mass vary by several orders of
magnitude inside a community, one downside of using abundance is that we mix
individuals that have potentially really different contributions in terms of ecosystem
properties such as total biomass and biomass repartition. We estimated
the body mass following an allometric law ($B_i = 0.01 \times
l_i^{3.03}$, $B$ being the body mass in grams, $l$ the body size in millimeters
and $i$ the individual) [@lleonart_removing_2000; @froese_cube_2006].

The species and total biomass were then computed for each sampling event.
As the sampling effort, reported as the surface sampled during the sampling event,
varied across protocol type and sampling event, biomass was divided by the
surface sampled (in $m^{2}$). So that, community and species biomass were expressed
in $gm^{-2}$.

In the analysis, the total biomass of a given station was characterized as the
median of the total biomass over time. Temporal biomass stability was defined as
the inverse of biomass variability. Variability was computed as Coefficient of
Variation (CV) of the biomass over time ($CV_i = \sigma_i \mu_i^{-1}$, with $\sigma_i$ and
$\mu_i$ being respectively the standard deviation and the temporal average of
the biomass of the station $i$). Then, temporal stability was equal to $CV^{-1}
= \mu \sigma^{-1}$.

To characterize the mechanisms driving temporal stability of communities, we
decomposed the community CV in synchrony and weighted average species CV
components according to @thibaut_understanding_2013:

$$
CV_{com, i} = \overline{CV_{sp, i}} \times \sqrt{\phi_i}
$$

with $CV_{com, i}$ being the Coefficient of Variation of the community in the
station $i$, $\overline{CV_{sp, i}}$ the average coefficient of variation of
population in the station $i$ pondered by the biomass of each population.
Synchrony of the station $i$, $\phi_i$, is the ratio between the variance of
total biomass over time ($\sigma^2_{x_T}$) and the sum of the variance of
population biomass over time ($\sigma_{x_k}$) [@loreau_species_2008]. $i$ being
the station $i$ and $k$ the species $k$.

$$
\overline{CV_{sp, i}} = \sum_k{\frac{B_k {CV}_i}{\sum_k{B_k}}}
$$

$$
\phi_i = \frac{\sigma^2_{x_T}}{(\sum_k{\sqrt{\sigma^2_{x_k}}})^2}
$$

## Horizontal & vertical diversity

@@

Species richness was computed as the total number of species present over time in a given
station. As for biomass, we controlled species richness for sampling effort by
dividing the total number of species by the total surface sampled in the given
station. Species richness was then expressed in number of species by square meter sampled.

@@

```{r}
myload(metaweb_analysis, dir= mypath("data"))
myload(trophic_class, dir= mypath("data", "classes"))
meta <- metaweb_analysis
#metaweb_analysis$resource
#meta$metaweb[meta$resource, meta$resource]
```

@@

For each trophic network, corresponding to a particular community at a given
time and location, we computed two of the most communly used network metrics to
characterize network structure, namely the connectance and average trophic
level. Connectance describes the number of interactions compared to the number
of possible links, computed as ($C = L / N^2$, L and N being resp. the number of
links and of trophic species plus resource nodes). The average trophic level was
computed as the mean trophic level of the nodes that the predator nodes consumes
to which is added one, obtained recursively by setting the trophic level of
basal nodes to 1, and was further weighted by the biomass of each trophic
species (excluding resource nodes for which we had no biomass information)
($\overline{T} = \sum_i^N{B_iT_i \times \sum_i^N{B_i}^{-1}}$, $T_i$ and $B_i$
being resp. the trophic level and the biomass of the node $i$). In resume, we
infered trophic networks at the scale of trophic species (i.e. nodes being body
size classes) to take in account ontogenic diet shifts and that fish-fish
trophic interaction depends on body size ratio between prey and predator. On the
other side, the community metrics were defined at the species level, (i.e.
species richness, average species CV and synchrony).

@@

For each fish station, environmental variables were matched with the year of
fish sampling (except for water temperature for which the record began in 2006).
Then, the average temperature, flow, BOD, river width and depth were computed as
well as coefficient of variation ($CV = \sigma \mu^{-1}$). As habitat and
environmental variables are often collinear, a principal component analysis (PCA)
was performed on the scaled variables. We kept in the analysis the two first
principal components based on eigen values distribution (Fig. SXX).
The PCA explained `r tot_var_expl_pca`%
of the variance. An varimax axis rotation was performed to maximize the representation
of the variables by the principal components (Fig. \@ref(fig:pca-rotated)).
The two first principal components were used in the analysis, representing respectively
the (1) river size (river width & depth), (2) altitude, slope and decreasing
average temperature. Further informations about data transformation and
filtering are detailed in supplementary information.

@@

@@

Structural Equation Modeling are able to assess multivariate relationships,
directs and indirects between variables. It is one promising way to go beyond
bivariate relationships and to investigate complex relationships, direct and
indirect, between variables [@grace_structural_2008; @grace_causal_2014;
@lefcheck_piecewisesem_2016; @grace_integrative_2016]. SEM makes possible to set
causal relationships between variables. This step can be critical, as the
estimated parameters can be determined by both the choice of the variables and
the links between them. Therefore, the set of variables and links should be
supported by theory and literature. We built a metamodel which present the set
of variables and the links between them (Figure \@ref(fig:matmet) A,
\@ref(fig:sem-plot) A & B). Basically, we tested the links between environmental
conditions, horizontal & vertical diversity and both total biomass and its
temporal stability (Figure \@ref(fig:matmet) A). We considered that environment
can directly affect horizontal and vertical diversity, as well as total biomass
and its temporal stability. Horizontal diversity, measured as species richness
directly affects vertical diversity, total biomass and its temporal stability.
Food web structure was shown to be constrained by species richness. In turn,
vertical diversity affects total biomass and its temporal stability (Figure
\@ref(fig:matmet) A). The metamodel describes how variables are interrelated
(Figure \@ref(fig:matmet) A). We built separated structural equation model for
total biomass and its temporal variability (Figure \@ref(fig:sem-plot) A & B).
